---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, message=F, warning=F }
library(dplyr)
library(lattice)
library(xtable)
```

## Loading and preprocessing the data


```{r}
activitydf<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
totalstepsdf<-filter(activitydf, !is.na(steps)) %>% group_by(date) %>% summarize(total_steps=sum(steps))
mean(totalstepsdf$total_steps)
median(totalstepsdf$total_steps)
```


```{r}
hist(totalstepsdf$total_steps, col="red", main="Total Steps per Day", xlab="Daily Number of Steps")
```


## What is the average daily activity pattern?

```{r}

mean_interval_stepsdf<-filter(activitydf, !is.na(steps)) %>% group_by(interval) %>% summarize(avg_steps=mean(steps))



plot(mean_interval_stepsdf$interval, mean_interval_stepsdf$avg_steps, type='l', xlab="Intervals", ylab="Avg Steps", main="Average Daily Pettern")

```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
mean_interval_stepsdf[which.max(mean_interval_stepsdf$avg_steps),]
```

##Calculate and report the total number of missing values in the dataset 

```{r}
length(activitydf[,1])-sum(complete.cases(activitydf))

```

###Devise a strategy for filling in all of the missing values in the dataset.

All the NA values appear in some specific days were we do not have any info.

```{r}

nadf<-filter(activitydf, is.na(steps))
xt<-(table(nadf$date))
print(xt, type="html")
```

For that reason i will assign for every NA the average number of steps

## Imputing missing values

```{r}

newdf<-activitydf

newdf[is.na(newdf[,1]),1]<-mean(activitydf$steps, na.rm=TRUE)

totalstepsnewdf<-select(newdf, 1:3) %>% group_by(date) %>% summarize(total_steps=sum(steps))


mean(totalstepsnewdf$total_steps)  
median(totalstepsnewdf$total_steps)

```

The mean is the same but the median changes sligthly

## Are there differences in activity patterns between weekdays and weekends?

```{r}


cleandf<-filter(activitydf, !is.na(steps))

cleandf$date<-weekdays(as.Date(cleandf$date, format = "%Y-%m-%d"))
cleandf$weekend<-cleandf$date
cleandf[which((cleandf[,2] %in% c("Saturday", "Sunday"))),4]<-"Weekend"

cleandf[which(!(cleandf[,2] %in% c("Saturday", "Sunday"))),4]<-"Weekday"

mean_interval_stepsdfv2<-select(cleandf, 1:4) %>% group_by(interval, weekend) %>% summarize(avg_steps=mean(steps))

mean_interval_stepsdfv2<-transform(mean_interval_stepsdfv2, weekend<-factor(weekend))
xyplot(avg_steps~interval | weekend, 
data=mean_interval_stepsdfv2, type='l', layout=c(1,2), ylab="Mean Steps per Interval")

```

From the plot we can see that there is differences in activity patterns between weekdays and weekends
